---
title: "Untitled"
editor: visual
---

```{r}
library(ggplot2)
library(readr)
library(here)
library(tidyverse)

```

```{r}

atp_singles_data <- readRDS(file = here :: here("data","processed","atp_singles.Rds" ))
```

The visualization below compares American players’ win percentages at the US Open and the Australian Open from 2000 to 2025, offering insight into potential home-court advantage over time. In the early 2000s, the US Open consistently favored American talent, with win rates hovering above 60% and peaking near 70%. However, that advantage noticeably eroded throughout the 2010s, with performance at both tournaments converging around the 40–50% mark as global competition intensified.

Interestingly, the past few years show a resurgence — particularly at the Australian Open — where American win percentages have climbed sharply again. This shift may reflect the emergence of a new generation of U.S. players who can succeed on multiple surfaces, reducing the historical reliance on hard-court familiarity and New York crowd support. Together, these trends suggest that while home-court advantage remains a factor in Grand Slam performance, American competitiveness is increasingly defined by adaptability and rising young talent capable of winning anywhere in the world.

```{r}

atp_singles_data %>%

  mutate(
    year = as.integer(format(as.Date(tourney_date), "%Y")),
    tn_clean = tourney_name |>
      str_to_lower() |>
      str_replace_all("\\.", "") |>
      str_squish(),
    tourney = case_when(
      str_detect(tn_clean, "^us\\s*open$") ~ "US Open",
      str_detect(tn_clean, "^australian\\s*open$") ~ "Australian Open",
      TRUE ~ NA_character_
    ),
    winner_ioc = toupper(winner_ioc),
    loser_ioc  = toupper(loser_ioc)
  ) %>%
  # filter relevant years, tournaments, and American matches
  filter(
    year >= 2000, year <= 2025,
    !is.na(tourney),
    winner_ioc == "USA" | loser_ioc == "USA"
  ) %>%
  mutate(american_win = (winner_ioc == "USA")) %>%
  group_by(tourney, year) %>%
  summarise(
    win_percent = 100 * mean(american_win, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # line graph instead of bar chart
  ggplot(aes(x = year, y = win_percent, color = tourney)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
 # geom_smooth(se = FALSE, method = "loess", linetype = "dashed", linewidth = 0.8) +
  labs(
    title = "American Player Win %: US Open vs Australian Open (2000–2025)",
    subtitle = "Visualizing home-court advantage over time",
    x = "Year",
    y = "Win Percentage (%)",
    color = "Tournament"
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

