---
title: "leading_country_final"
editor: visual
---


```{r}
#| message: false
#| warning: false
#| include: false
atp_singles_data <- readRDS(file = here :: here("data","processed","atp_singles.Rds" ))
```

The visualization below explores the historical match-win totals for five leading tennis nations, the United States, Spain, France, Italy, and Germany, spanning from 1968 to 2025. This broader international perspective highlights how different eras of tennis greatness have been shaped by the strength of national player cohorts. The United States stands out most prominently in the earlier decades, showing a dramatic peak in match wins throughout the 1970s and 1980s. This reflects a golden generation of American tennis led by icons such as Jimmy Connors, John McEnroe, and later Pete Sampras and Andre Agassi.

As we move toward the 2000s, a noticeable shift occurs in the competitive landscape. Match-win totals across countries become more balanced, signaling globalization of the sport and the rise of strong tennis cultures beyond the United States. Spain emerges as a new powerhouse, driven by clay-court depth and the sustained success of Rafael Nadal and other top Spanish players. France, Italy, and Germany all remain steady contenders, collectively contributing to an increasingly competitive field where success is distributed more evenly across nations.

This graph not only demonstrates the changing dominance in men’s tennis over time but also reflects how the rise of new generations and training infrastructures around the world continues to reshape the sport’s competitive dynamics.


```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(here)


selected_countries <- c("USA", "ESP", "FRA", "ITA", "GER")


graph2 <- wins_selected <- atp_singles_data %>%
  mutate(
    year = as.integer(format(as.Date(tourney_date), "%Y")),
    winner_ioc = toupper(str_trim(winner_ioc))
  ) %>%
  filter(
    !is.na(year),
    year >= 1968,
    winner_ioc %in% selected_countries
  ) %>%
  count(year, winner_ioc, name = "wins") %>%
  rename(country = winner_ioc)


ggplot(wins_selected, aes(x = year, y = wins, color = country)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.8) +
  labs(
    title = "Match Wins by Country Over Time ",
    subtitle = "USA, Spain, France, Italy, and Germany (1968–2025)",
    x = "Year",
    y = "Number of Match Wins",
    color = "Country"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

graph2
```
As the era of the “Big 3” comes to a close, the rise of young stars like Carlos Alcaraz and Jannik Sinner signals a new chapter in men’s tennis. With their rapid climb to the top of the rankings and early Grand Slam success, they appear poised to reshape the competitive landscape, continuing the trend of a more globally distributed field. Instead of a single dominant nation or trio of players, the future of tennis may be defined by tighter rivalries, more frequent shifts at the top, and increased representation from emerging tennis countries. This transition suggests a sport entering a more unpredictable and thrilling era, one where the next generation has the opportunity to redefine what sustained excellence looks like on the global stage.
